---
title: "Project1_EDA"
author: "Braden Anderson"
date: "1/31/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```



```{r}

source("Applied_Stats_Project_Functions.R")

df <- read.csv("./data1.csv", check.names=FALSE, na.strings=c("", "NA"))
names(df) <- gsub(" ", "_", names(df))

clean_df <- clean_data(df=df)
```



```{r}

#fullmodel = lm(logMSRP~Year+city.mpg+Popularity*Market.Category,auto)
#summary(fullmodel)
#plot(fullmodel)

```


# Investigate missing values
```{r, fig.width=10, fig.height=4}

# naniar::miss_case_table(df)

p1 <- plot_missings(df, 
                    omit_complete_columns = FALSE, 
                    plot_percent_missing = TRUE)

p2 <- plot_missings(df, 
                    omit_complete_columns = TRUE, 
                    plot_percent_missing = TRUE)



# PERCENTAGE of Missings before data cleaning
gridExtra::grid.arrange(p1, p2, ncol=2)
```

```{r, fig.width=10, fig.height=4}

p1 <- plot_missings(df, 
                    omit_complete_columns = FALSE, 
                    plot_percent_missing = FALSE)

p2 <- plot_missings(df, 
                    omit_complete_columns = TRUE, 
                    plot_percent_missing = FALSE)


# COUNT of Missings before data cleaning
gridExtra::grid.arrange(p1, p2, ncol=2)
```

# Missings after cleaning (currently rows with additional missings are being dropped, need to revisit this approach)
```{r, fig.width=10, fig.height=4}

p1 <- plot_missings(clean_df, 
                    omit_complete_columns = FALSE, 
                    plot_percent_missing = FALSE)

p1
```



# Investigate Correlations
```{r, fig.height=8, fight.width=8}

pearson_orig <- plot_correlation(df=df, 
                                 remove_columns=NULL, 
                                 corr_type="pearson", 
                                 title = paste("Pearson Correlation Matrix\nBefore Data Cleaning"), 
                                 hc_order=FALSE)



spearman_orig <- plot_correlation(df=df, 
                                  remove_columns=NULL, 
                                  corr_type="spearman", 
                                  title = paste("Spearman Correlation Matrix\nBefore Data Cleaning"), 
                                  hc_order=FALSE)


gridExtra::grid.arrange(pearson_orig, spearman_orig, ncol=2, newpage = TRUE)
```


```{r, fig.height=8, fight.width=8}

pearson_clean <- plot_correlation(df=clean_df, 
                                  col_names=NULL, 
                                  remove_columns=c("obs_number"), 
                                  corr_type="pearson", 
                                  title = paste("Pearson Correlation Matrix\nAfter Data Cleaning"),
                                  hc_order=FALSE  # Turning off smart ordering so all maps are ordered the same, for easier comparisons
                                  )


spearman_clean <- plot_correlation(df=clean_df, 
                                   col_names=NULL, 
                                   remove_columns=c("obs_number"), 
                                   corr_type="spearman", 
                                   title = paste("Spearman Correlation Matrix\nAfter Data Cleaning"), 
                                   hc_order=FALSE)

gridExtra::grid.arrange(pearson_clean, spearman_clean, ncol=2, newpage = TRUE)
```

# Attribute Data Types
```{r}

# Data types after data cleaning function
sapply(clean_df, class)

```


# Univariate Distributions
```{r}

# Target is right skewed
plot_histogram(df=clean_df, 
               continuous_variable="MSRP", 
               title="Distribution of MSRP", 
               binwidth=15000, 
               num_bins=NULL, 
               add_table=TRUE, 
               table_loc="upper_right")
```

```{r, fig.height=4, fig.width=12}

engine_hp_p <- plot_histogram(df=clean_df, 
                              continuous_variable="Engine_HP", 
                              binwidth=NULL, 
                              num_bins=NULL, 
                              add_table=TRUE, 
                              table_loc="upper_right", 
                              fill_color = "#FF69B4")

engine_cyl_p <- plot_histogram(df=clean_df, 
                               continuous_variable="Engine_Cylinders", 
                               binwidth=1, 
                               num_bins=NULL, 
                               add_table=TRUE, 
                               table_loc="upper_right")

gridExtra::grid.arrange(engine_hp_p, engine_cyl_p, nrow=1, ncol=2)
```



```{r, fig.height=4, fig.width=12}

highway_mpg_p <- plot_histogram(df=clean_df, 
                                continuous_variable="highway_MPG", 
                                binwidth=NULL, 
                                num_bins=NULL, 
                                table_loc="upper_right", 
                                fill_color="#00FFFF")

city_mpg_p <- plot_histogram(df=clean_df, 
                             continuous_variable="city_mpg", 
                             binwidth=NULL, 
                             num_bins=NULL, 
                             add_table=TRUE, 
                             table_loc="upper_right")

gridExtra::grid.arrange(highway_mpg_p, city_mpg_p, nrow=1, ncol=2)

```



```{r, fig.height=4, fig.width=12}

year_p <- plot_histogram(df=clean_df, 
                         continuous_variable="Year", 
                         binwidth=NULL, 
                         num_bins=NULL, 
                         table_loc="upper_left", 
                         fill_color="#39FF14")

doors_p <- plot_histogram(df=clean_df, 
                          continuous_variable="Number_of_Doors", 
                          binwidth=1, 
                          num_bins=NULL, 
                          add_table=TRUE, 
                          table_loc="upper_left")

gridExtra::grid.arrange(year_p, doors_p, nrow=1, ncol=2)
```


# Explore Log Transformations on right skewed features
```{r, fig.height=4, fig.width=12}

# Log transform seems to help some with the skew, but almost creats a bimodal distribution with a spike at 0.
trans_msrp <- plot_transform_comparison_hist(df=clean_df, 
                                             continuous_variable="MSRP", 
                                             fill_color_orig="#bc5090", 
                                             outline_color_orig="#003f5c", 
                                             fill_color_trans="#bc5090", 
                                             outline_color_trans="#003f5c", 
                                             num_bins_orig=NULL, 
                                             binwidth_orig=10000, 
                                             num_bins_trans=NULL, 
                                             binwidth_trans=0.2, 
                                             title_orig=NULL, 
                                             title_trans=NULL, 
                                             table_loc_orig="upper_right", 
                                             table_loc_trans="upper_right", 
                                             add_table_orig=TRUE, 
                                             add_table_trans=TRUE)
```



```{r, fig.height=4, fig.width=12}

trans_engine_hp <- plot_transform_comparison_hist(df=clean_df, 
                                             continuous_variable="Engine_HP")

```


```{r, fig.height=4, fig.width=12}

trans_highway_p <- plot_transform_comparison_hist(df=clean_df, 
                                                  continuous_variable="highway_MPG")

```


```{r, fig.height=4, fig.width=12}

trans_city_p <- plot_transform_comparison_hist(df=clean_df, 
                                               continuous_variable="city_mpg")

```


# Scatterplots of Continuous variables with Response (MSRP)
```{r, fig.height=4, fig.width=10}

p1 <- plot_scatter(data=clean_df, 
                   x_var="Engine_HP", 
                   y_var="MSRP", 
                   shade_var="Engine_Cylinders",   # Lets you shade points by a column in the dataframe
                   shape_var = NULL,               # Lets you change point shape by a column in the dataframe
                   size_var = NULL,                # Lets you change point size by a column in the dataframe
                   identify_obs = NULL,   # Observation numbers to identify. Useful when investigating influential points across multiple plots
                   alpha = 0.8,                    # Transparency of the markers
                   show_regression = FALSE,        # Should a regression be fit to these data?
                   reg_table = TRUE,               # If plotting a regression line, this lets you add a table with the regression coefs, test stat and p-value
                   filter_column="MSRP",           # Lets you specify a column to filter on prior to plotting, works in conjunction with the two params below
                   remove_less_than = NULL,        # Remove points in filter_column less than this value before potting
                   remove_greater_than = NULL,     # Remove points in filter_column greater than this value before plotting
                   remove_obs_numbers = NULL       # Allows you to remove points by observation number. You can get the observation numbers by setting identify_obs=TRUE
                   )
p1

```


```{r, fig.height=4, fig.width=15}

p1 <- plot_scatter(data=clean_df, 
                   x_var="MSRP", 
                   y_var="Popularity", 
                   shade_var="Engine_Cylinders",   
                   shape_var = NULL,               
                   size_var = NULL)

p2 <- plot_scatter(data=clean_df, 
                   x_var="Popularity", 
                   y_var="MSRP", 
                   shade_var="Engine_Cylinders",   
                   shape_var = NULL,               
                   size_var = NULL, 
                   remove_legend=FALSE)

gridExtra::grid.arrange(p1, p2, nrow=1, ncol=2)
```


```{r, fig.height=4, fig.width=15}

p1 <- plot_scatter(data=clean_df, 
                   x_var="MSRP", 
                   y_var="Popularity", 
                   shade_var="Year",   
                   shape_var = NULL,               
                   size_var = NULL)

p2 <- plot_scatter(data=clean_df, 
                   x_var="Popularity", 
                   y_var="MSRP", 
                   shade_var="Year",   
                   shape_var = NULL,               
                   size_var = NULL, 
                   remove_legend=FALSE)

gridExtra::grid.arrange(p1, p2, nrow=1, ncol=2)
```


```{r, fig.height=4, fig.width=10}

p1 <- plot_scatter(data=clean_df, 
                   x_var="Year", 
                   y_var="MSRP", 
                   shade_var="Engine_Cylinders",   # Lets you shade points by a column in the dataframe
                   shape_var = NULL,               # Lets you change point shape by a column in the dataframe
                   size_var = NULL,                # Lets you change point size by a column in the dataframe
                   identify_obs = NULL,   # Observation numbers to identify. Useful when investigating influential points across multiple plots
                   alpha = 0.8,                    # Transparency of the markers
                   show_regression = FALSE,        # Should a regression be fit to these data?
                   reg_table = TRUE,               # If plotting a regression line, this lets you add a table with the regression coefs, test stat and p-value
                   filter_column="MSRP",           # Lets you specify a column to filter on prior to plotting, works in conjunction with the two params below
                   remove_less_than = NULL,        # Remove points in filter_column less than this value before potting
                   remove_greater_than = NULL,     # Remove points in filter_column greater than this value before plotting
                   remove_obs_numbers = NULL       # Allows you to remove points by observation number. You can get the observation numbers by setting identify_obs=TRUE
                   )
p1

```

```{r}

clean_df[clean_df[,"Engine_Cylinders"] %in% c(8, 10, 12),]

```


```{r, fig.height=4, fig.width=10}

p1 <- plot_scatter(data=clean_df, 
                   x_var="Engine_Cylinders", 
                   y_var="MSRP", 
                   shade_var="Year",   # Lets you shade points by a column in the dataframe
                   shape_var = NULL,               # Lets you change point shape by a column in the dataframe
                   size_var = NULL,                # Lets you change point size by a column in the dataframe
                   identify_obs = NULL,   # Observation numbers to identify. Useful when investigating influential points across multiple plots
                   alpha = 0.8,                    # Transparency of the markers
                   show_regression = FALSE,        # Should a regression be fit to these data?
                   reg_table = TRUE,               # If plotting a regression line, this lets you add a table with the regression coefs, test stat and p-value
                   filter_column="MSRP",           # Lets you specify a column to filter on prior to plotting, works in conjunction with the two params below
                   remove_less_than = NULL,        # Remove points in filter_column less than this value before potting
                   remove_greater_than = NULL,     # Remove points in filter_column greater than this value before plotting
                   remove_obs_numbers = NULL       # Allows you to remove points by observation number. You can get the observation numbers by setting identify_obs=TRUE
                   )
p1

```


```{r, fig.height=4, fig.width=10}

p1 <- plot_scatter(data=clean_df, 
                   x_var="city_mpg", 
                   y_var="MSRP", 
                   shade_var="Engine_Cylinders",   # Lets you shade points by a column in the dataframe
                   shape_var = NULL,               # Lets you change point shape by a column in the dataframe
                   size_var = NULL,                # Lets you change point size by a column in the dataframe
                   identify_obs = NULL,   # Observation numbers to identify. Useful when investigating influential points across multiple plots
                   alpha = 0.8,                    # Transparency of the markers
                   show_regression = FALSE,        # Should a regression be fit to these data?
                   reg_table = TRUE,               # If plotting a regression line, this lets you add a table with the regression coefs, test stat and p-value
                   filter_column="MSRP",           # Lets you specify a column to filter on prior to plotting, works in conjunction with the two params below
                   remove_less_than = NULL,        # Remove points in filter_column less than this value before potting
                   remove_greater_than = NULL,     # Remove points in filter_column greater than this value before plotting
                   remove_obs_numbers = NULL       # Allows you to remove points by observation number. You can get the observation numbers by setting identify_obs=TRUE
                   )
p1

```

```{r, fig.height=4, fig.width=10}

p1 <- plot_scatter(data=clean_df, 
                   x_var="highway_MPG", 
                   y_var="MSRP", 
                   shade_var="Engine_Cylinders",   # Lets you shade points by a column in the dataframe
                   shape_var = NULL,               # Lets you change point shape by a column in the dataframe
                   size_var = NULL,                # Lets you change point size by a column in the dataframe
                   identify_obs = NULL,   # Observation numbers to identify. Useful when investigating influential points across multiple plots
                   alpha = 0.8,                    # Transparency of the markers
                   show_regression = FALSE,        # Should a regression be fit to these data?
                   reg_table = TRUE,               # If plotting a regression line, this lets you add a table with the regression coefs, test stat and p-value
                   filter_column="MSRP",           # Lets you specify a column to filter on prior to plotting, works in conjunction with the two params below
                   remove_less_than = NULL,        # Remove points in filter_column less than this value before potting
                   remove_greater_than = NULL,     # Remove points in filter_column greater than this value before plotting
                   remove_obs_numbers = NULL       # Allows you to remove points by observation number. You can get the observation numbers by setting identify_obs=TRUE
                   )
p1

```


# Factor Level Frequency Plots
```{r, fig.width=10, fight.height=5}

plot_factor_level_counts(df=clean_df,
                         factor_variable="Driven_Wheels")

```



```{r, fight.width=15}


plot_factor_level_counts(df=clean_df,
                         factor_variable="Year", 
                         xtick_rotation = 90, 
                         text_rotation=90,
                         txt_vjust=NULL, 
                         txt_hjust=1)
```

```{r, fig.width=12}

# TALK ABOUT THIS ONE

# Make
plot_factor_level_counts(df=clean_df,
                         factor_variable="Make", 
                         xtick_rotation = 90, 
                         text_rotation=90,
                         txt_vjust=NULL, 
                         txt_hjust=2)

```



```{r, fig.width=8}

# Vehicle Size
plot_factor_level_counts(df=clean_df,
                         factor_variable="Vehicle_Size")

```

```{r, fig.width=8}

plot_factor_level_counts(df=clean_df,
                         factor_variable="Engine_Cylinders")

```


```{r, fig.width=8}

# Number of doors
plot_factor_level_counts(df=clean_df,
                         factor_variable="Number_of_Doors")

```


```{r, fig.width=12}

# Vehicle Size
plot_factor_level_counts(df=clean_df,
                         factor_variable="Vehicle_Style", 
                         xtick_rotation = 90)

```


```{r, fig.width=12}

# TALK ABOUT THIS ONE

# Vehicle Size
plot_factor_level_counts(df=clean_df,
                         factor_variable="Transmission_Type", 
                         xtick_rotation = 90)

```


```{r, fig.width=12}


# Market Category
plot_factor_level_counts(df=clean_df,
                         factor_variable="Engine_Fuel_Type", 
                         xtick_rotation = 90, 
                         remove_legend=TRUE)


```


```{r}
clean_df[clean_df[,"Engine_Fuel_Type"] == "natural gas",]
```



```{r, fig.width=12}

# TALK ABOUT THIS ONE

# Market Category
plot_factor_level_counts(df=clean_df,
                         factor_variable="Market_Category", 
                         xtick_rotation = 90, 
                         remove_legend=TRUE)

```


```{r, fig.width=20}

# Model cannot be plotted due to the fact that there is 897 categories
# We can view the frequencies manually though (shown below)
#
# This column seems pretty much unusable. First because using this as categorical would be 897 regression 
# coefs to be fit, which is way to many coefs for this size of a data set. Second, because many of the 
# models only have a single instance, which would cause us issues when splitting into train and test sets... 
# If a car model shows up in the test set that was never seen during training, there will not be a regression coef 
# fit for that category, so the prediction will throw an error.
model_freqs <- data.frame(table(clean_df[,"Model"]))
model_freqs <- model_freqs[order(model_freqs[,"Freq"]),]
model_freqs <- model_freqs %>% rename(Model=Var1) 

# Run these for insight into why the Model column is pretty much unusable
head(model_freqs, n=10)
print(paste("Total Unique Models: ", nrow(model_freqs), " Total Models Only Represented Once: ", nrow(model_freqs[model_freqs[,"Freq"]==1,])))
```


# Categorical Mean Plots
```{r, fig.height=5, fight.width=15}

plot_means_by_category(df=clean_df, 
                       continuous_variable="MSRP", 
                       categorical_variable="Driven_Wheels")

```

```{r, fig.height=4, fight.width=15}

plot_means_by_category(df=clean_df, 
                       continuous_variable="MSRP", 
                       categorical_variable="Engine_Cylinders")

```


```{r, fig.height=5, fight.width=15}

plot_means_by_category(df=clean_df, 
                       continuous_variable="MSRP", 
                       categorical_variable="Transmission_Type")

```


```{r, fig.height=5, fight.width=15}

plot_means_by_category(df=clean_df, 
                       continuous_variable="MSRP", 
                       categorical_variable="Vehicle_Style")

```

```{r, fig.height=5, fight.width=15}

plot_means_by_category(df=clean_df, 
                       continuous_variable="MSRP", 
                       categorical_variable="Vehicle_Size")

```



```{r, fig.height=5, fight.width=20}

plot_means_by_category(df=clean_df, 
                       continuous_variable="MSRP", 
                       categorical_variable="Engine_Fuel_Type", 
                       xtick_rotation=90)

```

```{r, fig.height=4.5, fight.width=15}


# Make
plot_means_by_category(df=clean_df, 
                       continuous_variable="MSRP", 
                       categorical_variable="Make", 
                       xtick_rotation=90, 
                       remove_legend=TRUE)

```


```{r, fig.height=4.5, fig.width=15}

plot_means_by_category(df=clean_df, 
                       continuous_variable="MSRP", 
                       categorical_variable="Market_Category", 
                       remove_legend=TRUE,
                       xtick_rotation = 90)

```


```{r, fig.height=5, fight.width=17}

plot_means_by_category(df=clean_df, 
                       continuous_variable="MSRP", 
                       categorical_variable="Year")

```

```{r, fig.height=4, fight.width=17}

plot_means_by_category(df=clean_df, 
                       continuous_variable="MSRP", 
                       categorical_variable="Number_of_Doors")

```



# Splitting data into training, testing and validation sets

```{r}

source("Applied_Stats_Project_Functions.R")

df <- read.csv("./data1.csv", check.names=FALSE, na.strings=c("", "NA"))
names(df) <- gsub(" ", "_", names(df))

clean_df <- clean_data(df=df)
```


```{r}


#clean_df <- read.csv("./CLEAN_DATASET_0204.csv")

#write.csv(clean_df, "./CLEAN_DATASET_0204.csv", row.names=FALSE)

# Create the training, validation and test sets
split_datasets <- create_train_val_test_sets(df=clean_df, train_pct = 0.8, val_pct = 0.1)

train_df <- split_datasets$train
val_df <- split_datasets$val
test_df <- split_datasets$test


# Save the datasets to .csv for easy repeatability and sharing
write.csv(train_df, "./training_dataset_0208.csv", row.names=FALSE)
write.csv(val_df, "./validation_dataset_0208.csv", row.names=FALSE)
write.csv(test_df, "./testing_dataset_0208.csv", row.names=FALSE)


nrow(clean_df)
nrow(train_df)
nrow(val_df)
nrow(test_df)
```








