---
title: "Model_Evaluation"
author: "Braden Anderson"
date: "2/5/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


```{r}

source("Applied_Stats_Project_Functions.R")

# results_df <- read.csv("./model_checkpoints/all_models_0207.csv")

```

```{r}
# READ IN THE TRAIN, VALIDATION AND TEST SETS
train_df <- read.csv("./training_dataset_0206.csv")
val_df <- read.csv("./validation_dataset_0206.csv")
test_df <- read.csv("./testing_dataset_0206.csv")

# Ensure all datatypes are correct
train_df <- set_datatypes(train_df)
val_df <- set_datatypes(val_df)
test_df <- set_datatypes(test_df)
```


```{r}
summary(results_df)
```


```{r}

# Uncomment this and the file that reads in results_df if the consolidated best models file needs to be created
# best_unique_df <- get_best_models_df(metric_df=results_df)

# Read in the consolidated best models file
best_unique_df <- read.csv("./best_by_metric_consolidated.csv")

best_unique_df
```


# Best Model According to Validation MAE
```{r, fig.height=5, fig.width=15}

# Fit the model with the best metric_name value
fit_best_vmae <- get_best_model_by_metric(metric_df=best_unique_df, 
                                          training_df=train_df,
                                          metric_name="val_mae")


# Plot of extenernally studentized residuals
resid_plot_vmae <- plot_residuals(fit=fit_best_vmae, 
                                  dataframe=train_df, 
                                  id_extreme_values=TRUE)



# Plot of externally studentized residuals vs leverage
resid_lev_vmae <- plot_residual_vs_leverage(fit=fit_best_vmae, 
                                            dataframe=train_df)


gridExtra::grid.arrange(resid_plot_vmae, resid_lev_vmae, ncol=2)
```

# Normality Plots for Best Model According to Validation MAE
```{r, fig.height=5, fig.width=15}

# Normality Plots
resid_hist_vmae <- plot_residual_histogram(fit=fit_best_vmae,
                                           dataframe=train_df,
                                           binwidth=0.5, 
                                           overlay_normal=TRUE)

# QQ Plot of residuals
resid_qq_vmae <- plot_residual_qq(fit=fit_best_vmae, 
                                  dataframe=train_df, 
                                  flag_nlargest=6)

# Normality plots for model with best val_mae
gridExtra::grid.arrange(resid_hist_vmae, resid_qq_vmae, ncol=2)
```

# Case statistics plots for best model according to val_mae
```{r, fig.height=8, fig.width=15}

# Case statistics plots
cook_vmae <- plot_case_stat_vs_obs(fit=fit_best_vmae, dataframe=train_df, case_stat="cook", cook_rot=0.03)
leverage_vmae <- plot_case_stat_vs_obs(fit=fit_best_vmae, dataframe=train_df, case_stat="leverage", max_flagged=5)
resid_vmae <- plot_case_stat_vs_obs(fit=fit_best_vmae, dataframe=train_df, case_stat="externally_studentized_residuals", max_flagged=6)
del_std_vmae <- plot_case_stat_vs_obs(fit=fit_best_vmae, dataframe=train_df, case_stat="deleted_standard_deviation", std_rotm=0.01, max_flagged=6)
dffits_vmae <- plot_case_stat_vs_obs(fit=fit_best_vmae, dataframe=train_df, dffit_rotm=0.25, case_stat="dffit", max_flagged=6)

gridExtra::grid.arrange(cook_vmae, leverage_vmae, resid_vmae, dffits_vmae, ncol=2, nrow=2)
```


# Best Model per val_rmse and val_rsq
```{r, fig.height=5, fig.width=15}

# Fit the model with the best metric_name value
fit_best_vrmse <- get_best_model_by_metric(metric_df=best_unique_df, 
                                          training_df=train_df,
                                          metric_name="val_rmse")


# Plot of extenernally studentized residuals
resid_plot_vrmse <- plot_residuals(fit=fit_best_vrmse, 
                                  dataframe=train_df, 
                                  id_extreme_values=TRUE)



# Plot of externally studentized residuals vs leverage
resid_lev_vrmse <- plot_residual_vs_leverage(fit=fit_best_vrmse, 
                                            dataframe=train_df)




gridExtra::grid.arrange(resid_plot_vrmse, resid_lev_vrmse, ncol=2)
```


# Normality plots for best model per val_rmse and val_rsq
```{r, fig.height=5, fig.width=15}

# Normality Plots
resid_his_vrmse <- plot_residual_histogram(fit=fit_best_vrmse,
                                           dataframe=train_df,
                                           binwidth=0.5, 
                                           overlay_normal=TRUE)

# QQ Plot of residuals
resid_qq_vrmse <- plot_residual_qq(fit=fit_best_vrmse, 
                                  dataframe=train_df, 
                                  flag_nlargest=4)

# Normality plots for model with best val_mae
gridExtra::grid.arrange(resid_his_vrmse, resid_qq_vrmse, ncol=2)

```


# Case statistic plots for best model per val_rmse and val_rsq
```{r, fig.height=8, fig.width=15}

# Case statistics plots
cook_vrmse <- plot_case_stat_vs_obs(fit=fit_best_vrmse, dataframe=train_df, case_stat="cook", cook_rot=0.03)
leverage_vrmse <- plot_case_stat_vs_obs(fit=fit_best_vrmse, dataframe=train_df, case_stat="leverage", max_flagged=5)
resid_vrmse <- plot_case_stat_vs_obs(fit=fit_best_vrmse, dataframe=train_df, case_stat="externally_studentized_residuals", max_flagged=6)
del_std_vrmse <- plot_case_stat_vs_obs(fit=fit_best_vrmse, dataframe=train_df, case_stat="deleted_standard_deviation", std_rotm=0.01, max_flagged=6)
dffits_vrmse <- plot_case_stat_vs_obs(fit=fit_best_vrmse, dataframe=train_df, dffit_rotm=0.25, case_stat="dffit", max_flagged=6)


gridExtra::grid.arrange(cook_vrmse, leverage_vrmse, resid_vrmse, dffits_vrmse, ncol=2, nrow=2)
```



# Best Model per rmse, msep, rsq, adj_rsq
```{r, fig.height=5, fig.width=15}

# PREDICTORS USED:
# Year city_mpg Engine_Cylinders Engine_HP Transmission_Type Number_of_Doors 
# Vehicle_Style Vehicle_Size Driven_Wheels Factory_Tuner Performance Luxury Flex_
# Fuel Hatchback Hybrid Diesel Exotic Crossover Popularity

# Fit the model with the best metric_name value
fit_best_rmse <- get_best_model_by_metric(metric_df=best_unique_df, 
                                          training_df=train_df,
                                          metric_name="rmse")


# Plot of extenernally studentized residuals
resid_plot_rmse <- plot_residuals(fit=fit_best_rmse, 
                                  dataframe=train_df, 
                                  id_extreme_values=TRUE)



# Plot of externally studentized residuals vs leverage
resid_lev_rmse <- plot_residual_vs_leverage(fit=fit_best_rmse, 
                                            dataframe=train_df)



gridExtra::grid.arrange(resid_plot_rmse, resid_lev_rmse, ncol=2)
```


# Normality plots for best model per rmse, msep, rsq, adj_rsq
```{r, fig.height=5, fig.width=15}

# Normality Plots
resid_his_rmse <- plot_residual_histogram(fit=fit_best_rmse,
                                           dataframe=train_df,
                                           binwidth=0.5, 
                                           overlay_normal=TRUE)

# QQ Plot of residuals
resid_qq_rmse <- plot_residual_qq(fit=fit_best_rmse, 
                                  dataframe=train_df, 
                                  flag_nlargest=4)

# Normality plots for model with best val_mae
gridExtra::grid.arrange(resid_his_rmse, resid_qq_rmse, ncol=2)

```


# Case statistic plots for best model per rmse, msep, rsq, adj_rsq
```{r, fig.height=8, fig.width=15}

# Case statistics plots
cook_rmse <- plot_case_stat_vs_obs(fit=fit_best_rmse, dataframe=train_df, case_stat="cook", cook_rot=0.03)
leverage_rmse <- plot_case_stat_vs_obs(fit=fit_best_rmse, dataframe=train_df, case_stat="leverage", max_flagged=5)
resid_rmse <- plot_case_stat_vs_obs(fit=fit_best_rmse, dataframe=train_df, case_stat="externally_studentized_residuals", max_flagged=6)
del_std_rmse <- plot_case_stat_vs_obs(fit=fit_best_rmse, dataframe=train_df, case_stat="deleted_standard_deviation", std_rotm=0.01, max_flagged=6)
dffits_rmse <- plot_case_stat_vs_obs(fit=fit_best_rmse, dataframe=train_df, dffit_rotm=0.25, case_stat="dffit", max_flagged=6)


gridExtra::grid.arrange(cook_rmse, leverage_rmse, resid_rmse, dffits_rmse, ncol=2, nrow=2)
```



# Best Model per aic, apc, fpe, hsp, press, pred_rsq, train_mae
```{r, fig.height=5, fig.width=15}

# PREDICTORS USED:
# Year city_mpg Engine_Cylinders Engine_HP Transmission_Type Number_of_Doors 
# Vehicle_Style Vehicle_Size Driven_Wheels Factory_Tuner Performance Luxury 
# Flex_Fuel Hatchback Hybrid Diesel Exotic Popularity

# Fit the model with the best metric_name value
fit_best_aic <- get_best_model_by_metric(metric_df=best_unique_df, 
                                          training_df=train_df,
                                          metric_name="aic")


# Plot of extenernally studentized residuals
resid_plot_aic <- plot_residuals(fit=fit_best_aic, 
                                  dataframe=train_df, 
                                  id_extreme_values=TRUE)



# Plot of externally studentized residuals vs leverage
resid_lev_aic <- plot_residual_vs_leverage(fit=fit_best_aic, 
                                            dataframe=train_df)



gridExtra::grid.arrange(resid_plot_aic, resid_lev_aic, ncol=2)
```


# Normality plots for best model per aic, apc, fpe, hsp, press, pred_rsq, train_mae
```{r, fig.height=5, fig.width=15}

# Normality Plots
resid_his_aic <- plot_residual_histogram(fit=fit_best_aic,
                                           dataframe=train_df,
                                           binwidth=0.5, 
                                           overlay_normal=TRUE)

# QQ Plot of residuals
resid_qq_aic <- plot_residual_qq(fit=fit_best_aic, 
                                  dataframe=train_df, 
                                  flag_nlargest=4)

# Normality plots for model with best val_mae
gridExtra::grid.arrange(resid_his_aic, resid_qq_aic, ncol=2)

```


# Case statistic plots for best model per aic, apc, fpe, hsp, press, pred_rsq, train_mae
```{r, fig.height=8, fig.width=15}

# Case statistics plots
cook_aic <- plot_case_stat_vs_obs(fit=fit_best_aic, dataframe=train_df, case_stat="cook", cook_rot=0.03)
leverage_aic <- plot_case_stat_vs_obs(fit=fit_best_aic, dataframe=train_df, case_stat="leverage", max_flagged=5)
resid_aic <- plot_case_stat_vs_obs(fit=fit_best_aic, dataframe=train_df, case_stat="externally_studentized_residuals", max_flagged=6)
del_std_aic <- plot_case_stat_vs_obs(fit=fit_best_aic, dataframe=train_df, case_stat="deleted_standard_deviation", std_rotm=0.01, max_flagged=6)
dffits_aic <- plot_case_stat_vs_obs(fit=fit_best_aic, dataframe=train_df, dffit_rotm=0.25, case_stat="dffit", max_flagged=6)


gridExtra::grid.arrange(cook_aic, leverage_aic, resid_aic, dffits_aic, ncol=2, nrow=2)
```




# Best Model per sbc_SAS, sbc_R
```{r, fig.height=5, fig.width=15}

# PREDICTORS USED:
# Year Engine_Cylinders Engine_HP Transmission_Type Number_of_Doors 
# Vehicle_Style Vehicle_Size Driven_Wheels Factory_Tuner Performance Luxury 
# Flex_Fuel Hatchback Hybrid Diesel Exotic Popularity


# Fit the model with the best metric_name value
fit_best_sbc <- get_best_model_by_metric(metric_df=best_unique_df, 
                                          training_df=train_df,
                                          metric_name="sbc_SAS")


# Plot of extenernally studentized residuals
resid_plot_sbc <- plot_residuals(fit=fit_best_sbc, 
                                  dataframe=train_df, 
                                  id_extreme_values=TRUE)



# Plot of externally studentized residuals vs leverage
resid_lev_sbc <- plot_residual_vs_leverage(fit=fit_best_sbc, 
                                            dataframe=train_df)



gridExtra::grid.arrange(resid_plot_sbc, resid_lev_sbc, ncol=2)
```


# Normality plots for best model per sbc_SAS, sbc_R
```{r, fig.height=5, fig.width=15}

# Normality Plots
resid_his_sbc <- plot_residual_histogram(fit=fit_best_sbc,
                                           dataframe=train_df,
                                           binwidth=0.5, 
                                           overlay_normal=TRUE)

# QQ Plot of residuals
resid_qq_sbc <- plot_residual_qq(fit=fit_best_sbc, 
                                  dataframe=train_df, 
                                  flag_nlargest=4)

# Normality plots for model with best val_mae
gridExtra::grid.arrange(resid_his_sbc, resid_qq_sbc, ncol=2)

```


# Case statistic plots for best model per sbc_SAS, sbc_R
```{r, fig.height=8, fig.width=15}

# Case statistics plots
cook_sbc <- plot_case_stat_vs_obs(fit=fit_best_sbc, dataframe=train_df, case_stat="cook", cook_rot=0.03)
leverage_sbc <- plot_case_stat_vs_obs(fit=fit_best_sbc, dataframe=train_df, case_stat="leverage", max_flagged=5)
resid_sbc <- plot_case_stat_vs_obs(fit=fit_best_sbc, dataframe=train_df, case_stat="externally_studentized_residuals", max_flagged=6)
del_std_sbc <- plot_case_stat_vs_obs(fit=fit_best_sbc, dataframe=train_df, case_stat="deleted_standard_deviation", std_rotm=0.01, max_flagged=6)
dffits_sbc <- plot_case_stat_vs_obs(fit=fit_best_sbc, dataframe=train_df, dffit_rotm=0.25, case_stat="dffit", max_flagged=6)


gridExtra::grid.arrange(cook_sbc, leverage_sbc, resid_sbc, dffits_sbc, ncol=2, nrow=2)
```


# Best Model per rsq, train_rmse
```{r, fig.height=5, fig.width=15}

# PREDICTORS USED:
# Year highway_MPG city_mpg Engine_Cylinders Engine_HP Transmission_Type Number_of_Doors Vehicle_Style 
# Vehicle_Size Driven_Wheels Factory_Tuner Performance Luxury Flex_Fuel Hatchback Hybrid Diesel Exotic Crossover Popularity


# Fit the model with the best metric_name value
fit_best_trmse <- get_best_model_by_metric(metric_df=best_unique_df, 
                                          training_df=train_df,
                                          metric_name="train_rmse")


# Plot of extenernally studentized residuals
resid_plot_trmse <- plot_residuals(fit=fit_best_trmse, 
                                  dataframe=train_df, 
                                  id_extreme_values=TRUE)



# Plot of externally studentized residuals vs leverage
resid_lev_trmse <- plot_residual_vs_leverage(fit=fit_best_trmse, 
                                            dataframe=train_df)



gridExtra::grid.arrange(resid_plot_trmse, resid_lev_trmse, ncol=2)
```


# Normality plots for best model per rsq, train_rmse
```{r, fig.height=5, fig.width=15}

# Normality Plots
resid_his_trmse <- plot_residual_histogram(fit=fit_best_trmse,
                                           dataframe=train_df,
                                           binwidth=0.5, 
                                           overlay_normal=TRUE)

# QQ Plot of residuals
resid_qq_trmse <- plot_residual_qq(fit=fit_best_trmse, 
                                  dataframe=train_df, 
                                  flag_nlargest=4)

# Normality plots for model with best val_mae
gridExtra::grid.arrange(resid_his_trmse, resid_qq_trmse, ncol=2)

```


# Case statistic plots for best model per rsq, train_rmse
```{r, fig.height=8, fig.width=15}

# Case statistics plots
cook_trmse <- plot_case_stat_vs_obs(fit=fit_best_trmse, dataframe=train_df, case_stat="cook", cook_rot=0.03)
leverage_trmse <- plot_case_stat_vs_obs(fit=fit_best_trmse, dataframe=train_df, case_stat="leverage", max_flagged=5)
resid_trmse <- plot_case_stat_vs_obs(fit=fit_best_trmse, dataframe=train_df, case_stat="externally_studentized_residuals", max_flagged=6)
del_std_trmse <- plot_case_stat_vs_obs(fit=fit_best_trmse, dataframe=train_df, case_stat="deleted_standard_deviation", std_rotm=0.01, max_flagged=6)
dffits_trmse <- plot_case_stat_vs_obs(fit=fit_best_trmse, dataframe=train_df, dffit_rotm=0.25, case_stat="dffit", max_flagged=6)


gridExtra::grid.arrange(cook_trmse, leverage_trmse, resid_trmse, dffits_trmse, ncol=2, nrow=2)
```


# Investigate the Influential Points found in the above plots
```{r}

# High residual and low predicted price
train_df[(train_df[,"obs_number"] %in% c(1932, 1933, 1934, 5337, 5338, 5339, 5340, 5341, 5342)),]

```
```{r}

# High residual and high predicted price (not high leverage)
train_df[(train_df[,"obs_number"] %in% c(1369, 3464, 3934)),]

```


```{r}

# High residual and high predicted price (these also have high leverage)
train_df[(train_df[,"obs_number"] %in% c(1307, 5246, 5885)),]

```



```{r}

cat("first_line_test", file="./test_file.md")
cat("\n\n(./qqplot_test.png)", file="./test_file.md", append=TRUE)

ggsave(filename="qqplot_test.png",
       plot=p)

```

```{r}

#calculate_original_scale_metrics <- function(train_data, val_data, test_data, predictors_dataframe, filename){
#}

# Model Selection
# 1. Train models using training set
# 2. Estimate those models performance on new data / compare them, by predicting on the validation set
# 3. Pick a model
# 4. Test on test set final unbiased estimate of generalization error


# ------------------------------------------------------------

# Assess the Assumptions  (Normality, equal variance, independence, Linearity, Multicollinearity)
# 1. Residuals vs fitted values
# 2. Residual histograms
# 3. Residuals QQ Plots
# 4. Residuals vs Leverages 

# plot(fit)

# plot(x=predict(fit), y=train_df$log_MSRP,
#     xlab='Predicted Values',
#     ylab='Actual Values',
#     main='Predicted vs. Actual Values')
# summary(fit)
```


```{r}



```

```{r}


```


```{r}


```


```{r}


```


```{r}


```

```{r}


```


```{r}


```


```{r}


```


```{r}


```